---
title: "DOHC"
output: 
  bookdown::html_document2:
    base_format: rmarkdown::html_vignette
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{DOHC}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# Mapping Ocean Heat Content from Synthetic Observations in Indian Ocean

Knowledge of the spatio-temporal evolution of ocean heat content (OHC) is essential 
to our understanding of both past and future climate change. At global scales, ocean 
heat uptake is a key element of the global energy and sea level budgets and provides 
a powerful constraint on climate sensitivity and past climate forcings. At regional 
scales, OHC changes are important to our understanding of climate variability and 
for the development, initialization, and evaluation of seasonal-to-decadal prediction 
systems.

The world's oceans have a huge volume so even small changes in temperature can correspond 
to large changes in energy. 90% of the excess energy trapped in the earth system by 
anthropogenic greenhouse gas emissions goes into the oceans. 

## Measuring Ocean Heat Content

To measure heat content, we measure temperatures through the whole volume of the 
upper oceans, with measurements at different depths as well as at different locations. 

Historically, measurement of ocean temperature at depth have been obtained obtained using 
a range of instrument deployed from ships. More recently, since the early 2000s, 
autonomous robotic floats, known as Argo floats, have greatly increased the number 
of available observations of the upper 2000m of the world’s oceans. These data sources
provide profiles of ocean temperatures from the surface downwards. 

## Tutorial Data Source

This tutorial uses data from the MapEval4OceanHeat (ME4OH) project. The data
are derived from a 1/10th of a degree latitude-longitude ocean simulation that has been 
sub-sampled to real world observed locations. While the synthetic profile data are
intended to be used to benchmark ocean heat mapping methods, the data are used in this
tutorial as a case study for application of 4DModeller to map ocean heat content.

In this tutorial we will use  point data of ocean heat for upper layers of the ocean 
provided in the ME4OH data files (integrated temperature profiles from the surface down 
to a depth of 306.25 m, multiplied by water density and heat capacity).

The tutorial will use data from the Indian Ocean region. 

```{r}
library(INLA)
library(inlabru)
library(sf)
library(sp)
library(ggplot2)
```


## Loading Data
Here we load the Indian ocean shapefile. We smooth the boundary to make the mesh creation easier. 
```{r}
# Only for the first time 
# fdmr::retrieve_tutorial_data(dataset = "indian_ocean")
# fdmr::load_tutorial_data(dataset = "indian_ocean", 
#                          filename = "indian_ocean.rds")
ocean_sf <-
  st_read(here::here("OH","indian_ocean.shp"))
# ocean_sf <- ocean_sf[as.numeric(st_area(ocean_sf)) > 50000]
ocean_crs <- fm_crs(ocean_sf) 
fm_crs(ocean_sf) <- NULL
ocean_sf2 <- st_union(st_buffer(ocean_sf, 1.05))
```

We plot the Indian ocean to check if there is any issues. 
```{r}
ggplot() + geom_sf(data=ocean_sf2)

```

We convert into a sp object.
```{r}
fm_crs(ocean_sf) <- ocean_crs
ocean_sp <- as_Spatial(ocean_sf)
```

We regularly sample some points (20$\times$20) to put in the mesh builder function under the fdmr package to get a feeling of the parameter setting for the mesh.
```{r}
ocean_pts <- st_sample(ocean_sf, size = 20*20,
                       type = "regular")

ocean_pts_sp <- as_Spatial(ocean_pts)

colnames(ocean_pts_sp@coords) <- c("LONG", "LAT")
```

Here comes the mesh_builder function. We set the max edge and offset a bit larger to cater the computation. 
```{r}
fdmr::mesh_builder(spatial_data = as.data.frame(ocean_pts_sp),
                   max_edge = c(50,100),
                   offset = c(50,100),
                   # crs = fm_crs(ocean_pts_sp),
                   cutoff = 1)
```

We create the outer boundary. 
```{r}
ocean_bnd <- st_cast(
  st_sf(geometry = fm_nonconvex_hull(ocean_sf2, 5)),
  "POLYGON"
)
```

We create the mesh for the Indian ocean with the outer boundary.
We remove Sri Lanka and keep Madagasikara in the cutoff. We might do a Barrier model on a sphere in the future.  
```{r}
boundary <- fm_as_segm_list(list(ocean_sf2, ocean_bnd))

mesh <- fm_mesh_2d_inla(boundary = boundary,
                        max.edge = c(5, 7),
                        offset = c(3, 3),
                        cutoff = 1)

meshrds <- readRDS("mesh_5_7_1.rds")
me4 <- readRDS("me4_dohc.2005.indian_ocean.rds")
me4pts <- st_as_sf(me4, coords=c("lon","lat"))
ggplot() + geom_sf(data = ocean_sf2) + gg(mesh) + gg(data = me4pts, size=0.1)
```

## Model

We considered one year of top layer DOHC (between 0 and 306.25 m depth) in 2005. Observations are indexed by the month of occurrence, resulting in 12 indices, which is equal to the number of time points. To pass the data to the function, we created a `SpatialPolygonsDataFrame` from a `data.frame`.
```{r}
# Transform latitudes and longtiudes to numeric
df_f$lat <- base::sapply(df_f$lat, as.numeric)
df_f$lon <- base::sapply(df_f$lon, as.numeric)

# Add sp coordinates
sp::coordinates(df_f) <- c('lon', 'lat')

# Determine the group size
n_time <- base::as.integer(base::length(base::unique(df_f@data$time)))

```
### Prior selection

To implement the SPDE approach, we defined the range un uncertainty priors. We assume that the probability of process spatial range (physically, a distance where correlation between two observations falls to 0.1) being under 25 degrees in latitude is 0.2. We also set the probability to 0.1 that the marginal standard deviation of the process exceeds 1.

The process is assumed to be evolving temporally as an autoregressive process of the first order (AR1) with the probability of the temporal autocorrelation parameter be over 0 is 0.9.

```{r}
prior_range <- 25
prior_Pgt <- 0.2
  
# Define the SPDE
ohc_spde <- INLA::inla.spde2.pcmatern(mesh, prior.range = c(prior_range, prior_Pgt), prior.sigma = c(1, 0.1))

# Define temporal parameter
rhoprior <- base::list(theta = base::list(prior = 'pccor1', param = c(0, 0.9)))
```

### Define the model formula

We use a spatio-temporal model with only random effects.
```{r}
model_formula <- dohc_L1 ~ -1 + f(
  main = coordinates,
  model = ohc_spde,
  group = time,
  ngroup = n_time,
  control.group = list(
    model = 'ar1',
    hyper = rhoprior
  )
)
```

### Fit the model

```{r}
bru_model <- inlabru::bru(model_formula,
                          data = df_f,
                          family = "gaussian",
                          options = list(
                            verbose = TRUE,
                            bru_verbose = 4
                          )
)
```

## Output
```{r}
pred <- predict(bru_model)

pred_mean <- pred$Predictor["mean"]
pred_sd <- pred$Predictor["sd"]
mthpred <- list()
sdpred <- list()
for (i in 1:12) {
  mthpred[i]<-list(pred_mean[((1+1511*(i-1)):(1511*i)),])
  sdpred[i]<-list(pred_sd[((1+1511*(i-1)):(1511*i)),])
}

plot1 <- ggplot()+gg(mesh, color = as.numeric(unlist(mthpred[3]))) +  ggtitle("Mean March") + scale_fill_gradientn(name = "drhc (TJ/m^-2)",colours = c('navyblue', 'darkmagenta', 'darkorange1'), limits = c(0.33, 0.38)) + theme(legend.position="bottom", axis.title.x=element_blank(), axis.title.y=element_blank()) 
plot2 <- ggplot()+gg(mesh, color = as.numeric(unlist(mthpred[10]))) +  ggtitle("Mean October") + scale_fill_gradientn(name = "drhc (TJ/m^-2)",colours = c('navyblue', 'darkmagenta', 'darkorange1'), limits = c(0.33, 0.38)) +  theme(legend.position = "bottom", axis.title.x=element_blank(), axis.title.y=element_blank())
grid.arrange(plot1, plot2, ncol=2)  

plot1 <- ggplot()+gg(mesh, color = as.numeric(unlist(sdpred[3]))) +  ggtitle("Std March") + scale_fill_gradientn(name = "drhc (TJ/m^-2)",colours = c('navyblue', 'darkmagenta', 'darkorange1'), limits = c(0, 0.005)) + theme(legend.position="bottom", axis.title.x=element_blank(), axis.title.y=element_blank()) 
plot2 <- ggplot()+gg(mesh, color = as.numeric(unlist(sdpred[10]))) +  ggtitle("Std October") + scale_fill_gradientn(name = "drhc (TJ/m^-2)",colours = c('navyblue', 'darkmagenta', 'darkorange1'), limits = c(0, 0.005)) +  theme(legend.position = "bottom", axis.title.x=element_blank(), axis.title.y=element_blank())
grid.arrange(plot1, plot2, ncol=2)  


mean_mth <- lapply(mthpred, mean)
mean_std <- lapply(sdpred, mean)
plot1 <- ggplot(data = data.frame(x=1:12,y=unlist(mean_mth)), aes(x=x,y=y)) + geom_line() + labs(title="Mean over region ", x ="Month", y = "mean drhc (TJ/m^-2)") 
plot2 <- ggplot(data = data.frame(x=1:12,y=unlist(mean_std)), aes(x=x,y=y)) + geom_line() + labs( x ="Month", y = "std drhc (TJ/m^-2)") 
grid.arrange(plot1, plot2, nrow=2)  
```

